Building Markov State Models for Periodically Driven Non-Equilibrium Systems

### 2014-09-18 writedown/period.noneq.msm/paper.tex
CS's revise

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L^{\top}(t+T) \vect Y(t)=\vect L^{\top}(t)\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.
\recheck{This is known as the Floquet theorem~\cite{floquet1883equations}.}
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
and for the probability at integral period we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of the matrix $\myphi(T)$.

\recheck{
Using the Floquet theorem, the time-inhomogeneous Markov jump process $X_t$
is therefore discretized into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$, which is generated by transition matrix
\begin{equation}\label{P}
\vect P=\myphi^{\top}(T).
\end{equation}
We prefer to consider the discrete-time process $\tilde X_{m}$ instead of the time-continuous process $X_t$
because the powerful theories and
computational tools for time-homogeneous Markov processes can be directly applied. It worth noting that many of these tools require the transition matrix $P$ to satisfy the detailed balance condition. The computations in the Appendix~\ref{sec:app-revs} show that $P$ will in general \emph{not} satisfy this condition; in fact  the deviation form reversibility can be estimated from the work of the periodic driving does to the system.
There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behavior of the system on time-scales
much longer than the period will be perfectly described because of $\tilde X_{m} = X_{mT}\approx x_{mT}$ whenever our spatial discretization is fine enough.
At the same time,
the computational cost of generating $\tilde X_{m}$ is much less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.
}

Since \recheck{$P$} \recheck{(\underline{COMMENT}: using $P$ instead of $\myphi(T)$.)} is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with \recheck{right} eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a \recheck{left} eigenvector $\mymu$
satisfying
\recheck{$\mymu^{\top}P=\mymu^{\top}$.}
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). \recheck{In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.}

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov} the process $(\tilde X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\begin{align}
\label{eq:msm-bwd-committor}  
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j \mid X_m=i),
\end{align}
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
\recheck{\underline{QUESTION}: Here it is a \emph{conditional} probability,
  due to the time-homogeneity,
  it does not matter whether it is defined from quasi-stationary probability
  or not.}
From the last equation we get that the stationary distribution of the milestone process is given by
\begin{align}  
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\end{align}
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

\recheck{
We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ is associated with the milestone process $(\hat{X}_m)$ in terms of committors by}
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\recheck{
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$.

@@

\textbf{Remark 2:}
Following \cite{A19-31} we can also define another pair of stochastic MSM matrices: 
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j},
\end{align}
that are connected to $\hat{P}$ by the following identity
\[
\hat{P}=\hat{T}-\hat{M}+\id,
\]
That can be seen by means of direct computation: For the off-diagonal entries we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{jk} - \hat M_{jk}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \hat P_{jk}, \quad j\neq k
\end{align}
Stochasticity yields
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$ for the diagonal entries.  Furthermore, as shown in the Appendix~\ref{sec:app-prove}, $\hat{T}$ as well as $\hat{M}$ can be computed from trajectories without need to have the committors. 
The importance of the pair $\hat{T}$ and $\hat{M}$ for MSM building comes from the following observation: The main NEMD relaxation timescales are given by the dominant eigenvalues of $P$ \cite{A19-31, A19-1}.
These dominant eigenvalues can be approximated by discretizing the related eigenvalue problem $Pu=\lambda u$ by means of a Galerkin approximation with the finite dimensional ansatz space 
spanned by the forward committors $q^+_j$, $j=1,\ldots,K$ together  with the finite dimensional test function space spanned by the backward committors $q^-_j$, $j=1,\ldots,K$ (test functions multiplied from the left by the inner product $\langle\cdot,\cdot\rangle_\mu$). The thus discretize eigenproblem takes the form  of a generalized eigenproblem
\begin{align}
  \label{eq:msm-gen-ev}
\hat{T}\hat{u}=\hat{\lambda}\hat{M}\hat{u}.  
\end{align}
For the reversible case it is known that its $k$ eigenvalues $\hat{\lambda}$ are very good approximation of the dominant eigenvalues $\lambda$ of the original problem if the core sets are proper metastable sets~\cite{Eigenvalues}. Whether this is true for the non-reversible case is not known yet, but if the deviation from reversibility is weak and the dominant eigenvalues of $P$ are real-valued then the results should hold analogously, see~\cite{A19-31}, Thm.~4.19.

\textbf{Remark 3:}
The pair $\hat{T}$ and $\hat{M}$, or the matrix $\hat{T}\hat{M}^{-1}$ (if all of its entries are positive such that it is a stochastic matrix), thus can also be taken as MSM transition matrices. In the case of  $\hat{T}$ and $\hat{M}$ the MSM modelling error results from Galerkin discretization, while the MSM modelling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$.
\recheck{As a matter fact, if $\hat X_m$ is Markovian, the identity of the transition matrices holds: $\hat P = \hat T \hat M^{-1}$ (see~Appendix~\ref{sec:app-ptm} for the prove).
Therefore, in practice, the deviation of the eigenvalues of $\hat P$ from those of $\hat{T}\hat{M}^{-1}$ indicates the non-Markovianity implied in the process $\hat{X}_m$.}
The lesson learnt in the reversible case are that $\hat{P}$ is better suited to approximate the molecular kinetics on large timescales while $\hat{T}$ and $\hat{M}$ should be used for approximating eigenvalues and long relaxation timescales.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibirium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactoins,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of $10^6$~ps long, along which the snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^{\top}(T)$ just by counting the transition behavior of MD trajectories 
\[
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\]
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuos description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time \recheck{(small error in
  the finite difference approximation).}
\recheck{
  % COMMENT: shouldn't be discussions on the lag-time and the Markovianity?
  In the meanwhile,
  since the dimensionality is reduced by using only a few collective
  variables, the lag-time should be chosen large enough so that
  the original dynamics $x_t$ is properly relaxed at
  the unresolved degrees of freedoms in timescales shorter than the lag-time
  (possible if the collective variables capture the slow dynamics).
  Then the implied Markovianity assumption in the master equation is
  reasonable, and the discretization from $\ml(t)$ to $L(t)$ is state-independent.}
In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

\recheck{\underline{QUESTION}: Is the above discussion correct?
  Does the ``Markovianity'' always need a time discretization with the
  lag-time?
}

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{Time-dependent probability $\prob\big(\phi\in[0,180), \psi\in [0,180)\big)$.  The brute force NEMD simulation is compared with different
    spatial discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp2}
\end{figure}

>>>>

As an effect of the periodicity of $\vect L(t)$ the propagator $\myphi(t+T)$
satisfies
\begin{equation}\label{compo-1}
\myphi(t+T)=\myphi(t)\myphi(T),
\end{equation}
for all $t\ge 0$. This can be seen by considering $\vect Y(t)=\myphi(t+T)$. It satisfies
\[
\frac{d\ }{dt}\vect Y(t)=\vect L^{\top}(t+T) \vect Y(t)=\vect L^{\top}(t)\vect Y(t),\quad \vect Y(0)=\myphi(T).
\]
When we consider this identity column-wise and use the propagator property of $\myphi(t)$ we get $\myphi(t+T)=\vect Y(t)=\myphi(t)\myphi(T)$. As a consequence of (\ref{compo-1}) we get for all integers $m=0,1,2,\ldots$ that 
\begin{equation}\label{compo-2}
\myphi(t+mT)=\myphi(t)\myphi^m(T).
\end{equation}
In combination with Eq.~\eqref{eq:disc-8}, we therefore
know the solution $\vect p(t)$ of the Master equation for all $t\ge 0$,
if we can compute $\myphi(t)$ for $t\in (0,T)$.
This is known as the Floquet theorem~\cite{floquet1883equations}.
In particular we get the long-term evolution of the propagator:
\begin{align}
\label{eq:floq-13}  
\myphi(mT)=\myphi^m(T),
\end{align}
where $\myphi^m(T)$ denotes the $m$th power of $\myphi(T)$. Thus, for the probability at integral periods we have
\begin{align}
  \label{eq:floq-dynamics}
  p(mT) =  \myphi(mT)\, p(0) = \myphi^m(T)\, p(0).
\end{align}

Using the Floquet theorem, the time-inhomogeneous Markov jump process $X_t$
is therefore discretized into a \emph{time-homogeneous} (not necessarily
reversible) Markov jump process $\tilde X_{m} = X_{mT}, \ m\in\mathbb
N$, which is generated by transition matrix
\begin{equation}\label{P}
\vect P=\myphi^{\top}(T).
\end{equation}
We prefer to consider the discrete-time process $\tilde X_{m}$ instead of the time-continuous process $X_t$
because the powerful theories and
computational tools for time-homogeneous Markov processes can be directly applied. It worth noting that many of these tools require the transition matrix $P$ to satisfy the detailed balance condition. The computations in the Appendix~\ref{sec:app-revs} show that $P$ will in general \emph{not} satisfy this condition; in fact  the deviation form reversibility can be estimated from the work of the periodic driving does to the system.
There is no doubt that information within one period is lost by using this temporal
discretization, however, information regarding the long-term behavior of the system on time-scales
much longer than the period will be perfectly described because of $\tilde X_{m} = X_{mT}\approx x_{mT}$ whenever our spatial discretization is fine enough.
At the same time,
the computational cost of generating $\tilde X_{m}$ is much less demanding
than the brute force simulations of NEMD, which implies lower
statistically uncertainty in calculating the observables of interest.

Since $P$ is a stochastic matrix, its eigenvalues are contained in the unit circle in
the complex plane, i.e., each eigenvalue $\lambda$ (potentially coomplex-valued)
satisfies $|\lambda|\le 1$. Furthermore $\lambda=1$ is an
eigenvalue with right eigenvector $\vect e=(1,\ldots,1)^{\top}$ and a left eigenvector $\mymu$
satisfying
$\mymu^{\top}P=\mymu^{\top}$.
From now on, we assume $P$ to be irreducible and aperiodic such that the Perron Frobenius theorem the eigenvalue $\lambda=1$ is  non-negative componentwise, and unique (up to normalization $\sum_j\mu_j=1$). In this case $\mymu$ is the stationary measure in the sense that
$
 \mymu^{\top} P^m = \mymu^{\top},\ m\in\mathbb N
$,
% \begin{align}
% \label{eq:floq-14}  
% \end{align}
and (more precisely) the asymptotic evolution of an initial probability distribution $\vect p(t=0)$ by the process satisfies
$
\vect p^{\top}(0)\,P^m \to \mymu^{\top},\ m\to\infty,
$
% \begin{align}
% \label{eq:floq-15}  
% \end{align}
so that $\mymu$ can be seen as the quasi-stationary distribution of our non-stationary process.

@@

However, since we cannot assume $P$ to satisfy detailed balance, we
instead follow the approach to non-standard MSMs recently proposed in Ref.~\cite{sarich2014utilizing} which allows to identify the metastable core sets for the
non-reversible transition matrix $P$. Assume that this approach leads
to the $K$ core sets $C_1,\ldots, C_K\subset S$ that are appropriate metastable sets. 
Following~\cite{A19-31,schuette2011markov} the process $(\tilde X_m)$ associated with $P$ is coarse grained into the so-called milestone process $(\hat{X}_m)$ in the following way: 
\begin{itemize}
\item $\hat{X}_m$ just has $K$ states associated with the sets $C_j$, $j=1,\ldots,K$. 
\item The sequence of random variables $(\hat{X}_m)$ is defined via the sequence $(\tilde X_m)$, i.e., trajectories of $(\tilde X_m)$ induce trajectories of $(\hat{X}_m)$: We set $\hat{X}_m=j$ if the last core set that the process $(\tilde X_m)$ entered prior to or at time $m$ has been the core set $C_j$.
\end{itemize}
Now consider an arbitrary infinitely long trajectory of $(\tilde X_m)$. Because of ergodicity we know that the states in this trajectory will be distributed due to the quasi-stationary distribution $\mu$. Based on such an infinitely long trajectory we can consider the probability $q_j^\bwd(i)$ that conditioned on $\tilde X_m=i$ the last core set hit has been $C_j$.
This function is called the backward committor of $(\tilde X_m)$ associated with the set $C_j$ and is associated with the milestone process via
\begin{align}
\label{eq:msm-bwd-committor}  
q^\bwd_j(i)=\prob_\mu(\hat{X}_m=j \mid X_m=i),
\end{align}
where  the index $\mu$ refers to the fact that $\hat{X}_m$ is distributed due to $\mu$.
%\recheck{\underline{QUESTION}: Here it is a \emph{conditional} probability,
%  due to the time-homogeneity,
%  it does not matter whether it is defined from quasi-stationary probability
%  or not.}
From the last equation we get that the stationary distribution of the milestone process is given by
\begin{align}  
\hat{\mu}_j=\sum_{i\in S} q^\bwd_j(i)\mu(i),
\end{align}
that is, the probability to find $\hat{X}_m=j$ in (infinitely) long trajectories of the milestoning process is $\hat{\mu}_j$.

@@

We define the one-step transition matrix $\hat{\vect P}$ for the milestone process by
\begin{equation}\label{hatP_prob}
\hat{\vect P}_{jk}=\prob_\mu\Big(\hat{X}_{m+1}=k\,\vert\, \hat{X}_m=j\Big).
\end{equation}
Then Following
Ref.~\cite{djurdjevac2010markov}, Thm.~3.1, 
$\hat{\vect P}$ can be computed by matrix multiplication using the committors:
  \begin{align}
    \label{eq:msm-tmatrix-00}
    \begin{split}      
    \hat{\vect P}_{jk}
    = &
    \frac{1}{\hat{\mu}(j)}
    \langle (\vect P^b - \id) q^\bwd_j,q^\fwd_k \rangle_\mu,\qquad j\not= k, \\    %\label{eq:msm-tmatrix-01}
    \hat{\vect P}_{jj}
    =&
    1-\sum_{k\not=j} \hat{\vect P}_{jk}
    \end{split}
  \end{align}
where the inner product is defined by
$\langle u,v \rangle_\mu=\sum_{i\in S} u(i) v(i) \mu(i)$. 
In general the milestone process need not be a Markov process. The results in \cite{A19-31,sarich2014utilizing} show, however, that it is an approximate Markov process as long as the core sets are proper metastable sets, i.e., if the typical timecale on which $(X_m)$ leaves $C$ is much smaller than the typical expected hitting times between the core sets.
Thus, by taking $\hat{P}$ as our MSM transition matrix, we introduce an additional modelling error that is the smaller the more metastable the core sets are.
With this MSM transition matrix, we can define the MSM kinetics:
If we start from some initial probability $\hat{p}_j(0)$ of being in state $j$ at time $t=0$ then its evolution $\hat{p}_j(t)$ in time is discrete in multiples of period $T$ and given by
\begin{equation}\label{eq:num-29}
\hat{p}_j(T)=\sum_k \hat{p}_k(0)\hat{\vect P}_{kj}.
\end{equation}
In our approach $\hat p_j(mT)$ is a good approximation of $\prob(\hat{X}_m=j)$ (for appropriately chosen core sets).

@@

\textbf{Remark 2:}
Following \cite{A19-31} we can also define another pair of stochastic MSM matrices: 
\begin{align}\label{eq:msm-ht-1}
  \hat T_{jk} &= \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j},\\\label{eq:msm-hm-1}
  \hat M_{jk} &= \frac{\langle q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j},
\end{align}
that are connected to $\hat{P}$ by the following identity
\[
\hat{P}=\hat{T}-\hat{M}+\id,
\]
That can be seen by means of direct computation: For the off-diagonal entries we have
\begin{align}\label{eq:msm-tmp26}
  \hat T_{jk} - \hat M_{jk}
  = \frac{\langle q_j^\bwd, P q_k^\fwd \rangle_\mu}{\hat \mu_j}
  - \frac{\langle q_j^\bwd,  q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \frac{\langle (P^b - \id) q_j^\bwd, q_k^\fwd \rangle_\mu}{\hat \mu_j}
  = \hat P_{jk}, \quad j\neq k
\end{align}
Stochasticity yields
$\hat T_{jj} - \hat M_{jj} + 1 = \hat P_{jj}$ for the diagonal entries.  Furthermore, as shown in the Appendix~\ref{sec:app-prove}, $\hat{T}$ as well as $\hat{M}$ can be computed from trajectories without need to have the committors. 
The importance of the pair $\hat{T}$ and $\hat{M}$ for MSM building comes from the following observation: The main NEMD relaxation timescales are given by the dominant eigenvalues of $P$ \cite{A19-31, A19-1}.
These dominant eigenvalues can be approximated by discretizing the related eigenvalue problem $Pu=\lambda u$ by means of a Galerkin approximation with the finite dimensional ansatz space 
spanned by the forward committors $q^+_j$, $j=1,\ldots,K$ together  with the finite dimensional test function space spanned by the backward committors $q^-_j$, $j=1,\ldots,K$ (test functions multiplied from the left by the inner product $\langle\cdot,\cdot\rangle_\mu$). The thus discretize eigenproblem takes the form  of a generalized eigenproblem
\begin{align}
  \label{eq:msm-gen-ev}
\hat{T}\hat{u}=\hat{\lambda}\hat{M}\hat{u},\quad\mathrm{or, equivalently}\quad \hat{M}^{-1}\hat{T}\hat{u}=\hat{\lambda}\hat{u},.  
\end{align}
For the reversible case it is known that its $k$ eigenvalues $\hat{\lambda}$ are very good approximation of the dominant eigenvalues $\lambda$ of the original problem if the core sets are proper metastable sets~\cite{Eigenvalues}. Whether this is true for the non-reversible case is not known yet, but if the deviation from reversibility is weak and the dominant eigenvalues of $P$ are real-valued then the results should hold analogously, see~\cite{A19-31}, Thm.~4.19.

If all of its entries are positive such that it is a stochastic matrix, $\hat{M}^{-1}\hat{T}$ thus can also be taken as MSM transition matrices. In the case of $\hat{M}^{-1}\hat{T}$ the MSM modelling error results from Galerkin discretization, while the MSM modelling error of $\hat{P}$ results from ignoring the potential non-Markovianity of $\hat{X}_m$.

\textbf{Remark 3:} As a matter of fact, if $\hat X_m$ were Markovian, the following identity would hold: $\hat P = \hat T \hat M^{-1}$ (see~Appendix~\ref{sec:app-ptm} for the proof). In this case, we would have $\hat{M}^{-1}\hat{T}=\hat{M}^{-1}\hat{P}\hat{M}$ and the eigenvalues of $\hat{P}$ and $\hat{M}^{-1}\hat{T}$ would be identical.
Therefore, in practice, the deviation of the eigenvalues of $\hat P$ from those of $\hat{M}^{-1}\hat{T}$ indicates the deviation from Markovianity regarding the process $\hat{X}_m$.
%The lesson learnt in the reversible case are that $\hat{P}$ is better suited to approximate the molecular kinetics on large timescales while $\hat{M}^{-1}\hat{T}$ should be used for approximating eigenvalues and long relaxation timescales.

@@

To illustrate how the discretization works in practice we take the alanine dipeptide system under an oscillatory EF,
as an example, the NEMD simulation of which was
extensively studied in Ref.~\cite{wang2014exploring}.
The system was simulated in a $2.7\times 2.7\times 2.7$~$\textrm{nm}^3$ periodic simulation
region, with one alanine dipeptide molecules described by the CHARMM27 force field~\cite{foloppe2000all} dissolved in 641 TIP3P water molecules~\cite{jorgensen1983comparison}.
The grid-based energy correction map (CMAP)~\cite{mackerell2004extending} was used
to correct the backbone dihedral angle energies.
All simulations were performed by a home-modified
Gromacs~4.6.5~\cite{pronk2013gromacs} with CHARMM27 force field implemented~\cite{bjelkmar2010implementation}.
The alanine dipeptide was put into the local thermostating
environment, with a spherical dynamical region of radius~1.0~nm
centered at the alpha-carbon.
The Langevin thermostat with target temperature $\mathcal T = 300$~K
and time-scale $\tau_T = 0.1$~ps was
coupled to the thermostated region.
The whole system was coupled to the Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} (in standard Gromacs implementation) with $\tau_P = 2.0$~ps to
keep the system at 1~Bar. The non-equilibirium trajectories
were integrated by the Leap-frog scheme with a time-step of 0.002~ps.
The short-range van der Waals interactions were cut-off at 1.00~nm, and were smoothed from
0.95~nm to 1.00~nm by the ``\texttt{shift}'' method provide by Gromacs.
The energy conserving 
Particle Mesh Ewald (PME)~\cite{darden1993pme, essmann1995spm} method (``\texttt{pme-switch}'') was
used to compute the long-range electrostatic interaction,
with the same real-space cut-off radius as the van der Waals interactoins,
and the Gromacs default Fourier spacing of 0.12~nm.
The splitting
parameter optimized with respect to the electrostatic force computing
accuracy by Gromacs tool \texttt{g\char`_pme\char`_error}~\cite{wang2010optimizing}.
The neighbor list was updated every 5 time-steps with a list-building radius 1.20~nm.
All hydrogen involving covalent bonds were constrained by the LINCS algorithm~\cite{hess1997lincs}, except the water molecules that were constrained by the SETTLE algorithm~\cite{miyamoto2004settle}.
The whole system was driven by a periodic electric field
$E(t) = E_0\sin(2\pi t/T)$ and $D(x) = (1,0,0)^{\top}$
with intensity of the field being $E_0 = 1.0$~V/nm and period being
$T = 10$~ps.
The 20,000 branching trajectories were simulated from 20,000
initial configurations that sample the equilibrium distribution.
% \redc{Write a lot of details on NEMD.}
The equilibrium configurations were prepared by an equilibrium MD simulation
of length $10^6$~ps, along which snapshots were saved every 50~ps.
% In the
% equilibrium simulation the system was coupled to a global Langevin thermostat
% with coupling time-scale $\tau_T = 0.5$~ps.
The branching NEMD
trajectories were each 4000~ps long, and the system reached
non-equilibrium quasi-stationary state in roughly 300~ps.

@@

Based on a given spatial discretization we can aggregate the transition matrix $\vect P=\myphi^{\top}(T)$ just by counting the transition behavior of MD trajectories 
\begin{equation}\label{Pdirect}
\vect P_{ij}=\prob\left(\vect X_T=j\mid \vect X_0=i \right).
\end{equation}
However, $\vect P$ allows to approximate the original dynamics on multiples $mT$ of the period only. In order to have a time-continuos description we need the generator $L(t)$ of the Master equation. 
If the discretization is fine enough one possible approximation to $L(t)$ is via
the following forward finite difference scheme:
\begin{align}
  \label{eqn:tmp4}
  L_{ij}(t) \approx \frac{1}{\tau}
  \,[\, \prob (\vect X_{t+\tau} = j \mid \vect X_{t} = i) - \delta_{ji} \,],
  \quad i,j\in S
\end{align}
and $\tau$ is an appropriate small enough lag-time.
  Since the dimensionality is reduced by using only a few collective
  variables, the lag-time should be chosen large enough so that
  the original dynamics $x_t$ is properly relaxed with regard to
  the unresolved degrees of freedoms on timescales shorter than the lag-time
  (assuming  that the collective variables capture the slow dynamics).
  In the following 
we investigate the discretization
quality with respect to  the choice of $N_\dih$ and  lag-time $\tau$.

%\recheck{\underline{QUESTION}: Is the above discussion correct?
%  Does the ``Markovianity'' always need a time discretization with the
%  lag-time?
%}
%

\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{fig/t010/discrete/fig-cg-prob.eps}  
  \caption{Time-dependent probability $\prob\big(\phi_t\in[0,180), \psi_t\in [0,180)\big)$.  The brute force NEMD simulation is compared with different
    spatial discretization methods. The red shadow region indicates the
    statistical uncertainty of the NEMD simulation.}
  \label{fig:tmp2}
\end{figure}
